\section{Thesis Summary}


This Thesis reports algorithms for state estimation, adaptive
parameter identification, and model-based control  principally for
\ac{UV} applications.
%
An analytical proof of stability is included for every reported algorithm.
%
Chapter \ref{chSMS_ID} reports an angular velocity observer for
rotating rigid-bodies, an \ac{AID} algorithm for rotating rigid-bodies,
and an \ac{AID} algorithm for \acp{OKC}.
%
Numerical simulations of the \ac{AID} algorithm for rotating
rigid-bodies corroborate the analytical stability analysis and
investigate parameter convergence with varying initial conditions,
adaptation gains, and input torques.
%
Chapter \ref{chSMS_ID} also reports a comparative analysis of three
nonlinear angular velocity observer for rotating rigid-bodies; in
numeric simulations the novel angular velocity observer (Theorem
\ref{chSMS_ID.theo.AVO_Theorem}) is shown to provide performance
similar to that of the two previously reported observers.


Chapter \ref{chUV_AID} reports two \ac{AID} algorithms for the dynamic
estimation of \ac{UV} plant parameters.
%
Chapter \ref{chUV_AID} also reports two comparative experimental
evaluations of \acl{AID} and \acl{LS}.
%
Both the \acp{AIDPM} and the \acp{LSPM} are shown to match closely the
experimentally observed \ac{UV} input-output behavior.
%
Adaptive identification algorithms do not require simultaneous
reference trajectory-tracking control, nor do they require
instrumentation of linear acceleration or angular acceleration.
%
Together, these facts make adaptive identification applicable to a
wider class of \acp{UV} than previously reported methods.
%
Chapter \ref{chUV_AMBC} reports an \ac{UV} \ac{MBC} algorithm and
\ac{UV} \ac{AMBC} algorithm.  
%
Chapter \ref{chUV_AMBC} also reports an experimental evaluation of the
destabilizing effects of unmodeled thruster dynamics on \ac{AMBC} and a
two-step \ac{AMBC} algorithm which is shown experimentally to be
robust in the presence of unmodeled thruster dynamics.
%
A comparative experimental analysis of the two-step \ac{AMBC} and
\ac{PDC} is reported; it showed that \ac{AMBC} provides 30\% better
position tracking performance and marginally worse (8\%) velocity
tracking performance over \ac{PDC}.
