It is a simple truth that I could not have made it this far without
others investing in me.
%
My education was supported by the University of Puget Sound through
several scholarships; Washington University in St. Louis through a
Browns Fellowship for dual degree engineering students; and the Johns
Hopkins University Mechanical Engineering Department through a
Departmental Fellowship.
%
My education and research were supported by the Department of Defense,
Air Force Office of Scientific Research through a National Defense
Science and Engineering Graduate Fellowship (32 CFR 168a); the
National Science Foundation through awards 812138 and 1319667; the
Link Foundation through a Link Doctoral Research Fellowship in Ocean
Engineering and Instrumentation; and the ARCS Foundation through their
fellowship for science and technology scholars.



This Thesis is dedicated to my teachers, i.e.  those who have invested
effort in helping me become a better scientist, engineer, and person.
%
Teachers can be professors, friends, family members, colleagues, in
addition to the ``teachers'' who guided my childhood education.
%
I am proud of the contributions contained herein; however, when I look
back on all the effort others have invested in me, this work seems
inadequate.



I came to Johns Hopkins University for one reason.  I thought
Dr.\ Whitcomb was exactly the type of engineer I wanted to be.
%
Dr.\ Whitcomb, I was correct in that belief and am grateful you have
devoted so much of yourself to helping me become an engineer like you.
%
Over the past decade I have had several mentors who were always ready
to discuss new research directions over the phone or in-person.
%
Dr.\ Yoerger, Dr.\ Cowan, Dr.\ Kobilarov, Dr.\ Sarma, Dr.\ Smart, Mr. Few,
and Mr. Cain, I have enjoyed moments of friendship and scholarship
with each of you.  
%
Thank you for taking these opportunities to teach me the difficult art
of picking what problems are worth attempting to solve (and the many,
many letters of recommendation).
%
I am especially grateful for Dr.\ Yoerger, Dr.\ Cowan, Dr.\ Kobilarov,
and Dr.\ Whitcomb taking the time to serve on my Thesis committee and
improve the work contained herein.


So many educators have been instrumental in getting me here by
immersing me in physics, engineering, and the arts.
%
Dr.\ Thorndike, Dr.\ Elliott, Dr.\ Weber, Dr.\ Beezer, 
%
Dr.\ Stockdale, 
%
Dr.\ Tarraf, Dr.\ Fishkind, 
%
Dr.\ Genin,
%
Mr. Calkins, Dr.\ Jones, Mr. Dyer, Mr. Shen, Mrs. Holmes, Mr. Sweeney,
Mr. Palma,
%
Mr. Johnson, 
%
Mrs. Johnson, Mrs. Barnett, Mrs. Brown, and  
%
those I have forgotten to mention, 
%
words can not express how grateful I am for the countless hours spent
helping me better understand the world.  


Over the past 6 years I have been surrounded by so many fantastic
colleagues.  
%
Steve, Kevin, Mike, Tricia, John, Axel, Tom, Carol, Netta, David,
Tiffany, Hunter, Tomo, Kirk, Claire, Adrien, Manu, Kel, Xingchi,
Yixin, and Zak, thanks for always being willing to teach me
something new, just talk about research, or have a fun night out.
%
During these years an amazing group of close friends have poured
themselves into helping me become who I am today.
%
Drew, 
%Madeline, Jason, 
Grant, Karrie, Pat, 
%
%Good Jon, 
Marisa, 
%Galen, Travis, Evil John, Emily, Casey, 
%
%Jake, Joe, 
Lauren, 
%
Giancarlo, Amy, Sarah, 
%Steve, Axel,
Marcin, 
%Tom, 
Eatai,
% John, David, Tiffany, 
%
%Hunter, 
Maya,
%
%Tomo, Kirk, 
Jon, Chad, Jowita, and Yan Yan each of you deserve a
paragraph inserted here.  
%
In its stead, let me offer a heartfelt (if inadequate) thank you for
your love and support.


In closing, I would like to acknowledge my parents, the most important
teachers I have ever had.
%
I am so grateful for your enduring education in how to lead a happy,
successful life.
%
Thank you.