%\section{Introduction}
%\label{chModels.sec.intro}

This Chapter summarizes the notation, kinematic conventions, and plant
models used in this Thesis.
%
With the exception of Section \ref{chModels.sec.genMat_Factorization},
the models, functions, state spaces, and facts stated herein are commonly
employed in the fields of geometric control,
adaptive control, and  underwater robotics. 
%
Section \ref{chModels.sec.notation} states the functions, norms, and
\ac{SPD} matrix eigenvalue conventions used herein.
%
Section \ref{chModels.sec.genMat_Factorization} reports a proof of
Theorem \ref{chModels.theo.diagMat_Factorization}, which allows
general 3-by-3 matrices to be factored through the skew symmetric
operator.
%
Sections \ref{chModels.sec.SO3} and \ref{chModels.sec.SE3} detail
the position and velocity state representations used herein for a
rotating rigid-body and general rigid-body motion.
%
The analytic relationships between these state representations are
also stated.
%
This Chapter presents second-order models for:
\begin{itemize}
\item 3-\ac{DOF} rotating rigid-body dynamics (Section \ref{chModels.sec.SO3plant}), 
\item 3-\ac{DOF} rotating \ac{UV} dynamics  (Section \ref{chModels.sec.UVSO3plant}), 
\item 6-\ac{DOF} \ac{UV} dynamics (Section \ref{chModels.sec.UVSE3plant}), and 
\item {\it n}-link \ac{OKC} dynamics (Section \ref{chModels.sec.OKC}).
\end{itemize}
%
For each of the four models, plant parameter properties and the
model's regressor matrix factored form are also presented.
%
Readers seeking a presentation of this material in greater detail are
referred to the texts cited in Section \ref{chModels.sec.litReview}.


As noted in the summery above, we believe Theorem
\ref{chModels.theo.diagMat_Factorization} is a novel result.
%
This Theorem states that $A^T((Ax)\times(Ay))=\det(A)(x\times y)$ for
all $A\in\rSp{3\times 3}$ and $x,y\in\rSp{3}$, where $\times$ is the
usual cross product operator for $\rSp{3}$ and $\det(\cdot)$ is the
matrix determinate.
%
This result is a generalization of the well known fact $(Rx)\times y=
R(x\times(Ry))$ for all $x,y\in\rSp{3}$ and $R\in\SO3$, where SO(3) is
the special orthogonal group for $\rSp(3)$.
%
To the best of our knowledge, this relation has not been proposed or
proven previously.


