\section{ Adaptive Identification for 3-\acs{DOF} Rotational Plants 
  }\label{chSMS_ID.sec.SO3_AID}

This Section addresses the problem of estimation the inertia tensor
of 3-\ac{DOF} rotational plants of the form
(\ref{chModels.eq.SO3plant}).
%
We report a novel \ac{AID} algorithm which estimates the inertia
tensor for a rotating rigid-body using the signals of external torque
and angular velocity.
%
A local stability proof of the new \ac{AID} algorithm is reported.
% 
A numerical simulation study investigates the effect of richness of
the torque input signal on parameter estimation, the effect of
feedback gains, the effect of initial condition, and the domain of
stability.
%
The simulation studies corroborate the analytic stability analysis,
showing that the angular velocity estimate converges asymptotically to
the actual angular velocity of the plant and the adaptive estimate of
the inertia tensor converges to a constant value.
%
Additionally, the simulation studies show that the inertia tensor
estimate converges to the true plant inertia tensor value in the
presence of a sufficiently rich input torque signal.
%
The simulation studies reveal the actual domain of attraction to
exceed the conservative bounds arising in the stability proof, and
identify practically useful ranges of feedback gains.


\input{chSMS_ID/SO3_AID_Der}
\input{chSMS_ID/SO3_AID_Sim}

\subsection{ 3-\acs{DOF} Rotational Plant \acs{AID} Conclusion}

This Section reports an \ac{AID} algorithm for the dynamic estimation
of the inertia tensor of rotating plants.
%
The proof of Theorem \ref{chSMS_ID.theo.SO3_AID} shows the local
asymptotic stability of the estimated angular velocity to the plant's
angular velocity and the local stability of the estimated inertia
tensor. 
%
Numerical simulations show that for a sufficiently rich external
torque signal the inertia tensor estimate value converges to the true
inertia tensor value, and the domain of attraction of Theorem
\ref{chSMS_ID.theo.SO3_AID} is conservative.
%
In Chapter \ref{chUV_AID}, Theorems \ref{chUV_AID.theo.UV_SO3_AID} and
\ref{chUV_AID.theo.UV_SE3_AID} extend this result to plant models of
\ac{UV} dynamics.
